### 2015-05-20 reports/paper/experiment.tex
paper update

\subsubsection{Movement tracking}
To test how well the arrays track movement, we mounted a rotating disk $40$ centimeter in diameter onto the grid at $(x=0, y=-0.3)$. Fig~\ref{fig:setup_circle} shows a picture of the sestup. A sound source is placed on the edge of rotating disk and the arrays localizes the sound source as it rotates in a circle. In this experiment, we tested how accuracy changes with:
\begin{itemize}
\item different window sizes
\item different audio sources
\item different movement tracking filters
\item different movement speeds
\end{itemize}

@@

\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. "Honest Eyes" by Black Tide was the music used.

\item[Music B] A music with intermittent low amplitude sections was chosen. "Canon" was the music used.

@@

To test how accuracy varies with window size, the algorithm is fed with recorded microphone data with different segment length. Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with window length for three GCC algorithms. The error lowers as window size increases and plataeus after window size exceeds around $10$ milli-second. The lowest error achieved is $2.53$ centimeters. It is achieved when window size is set to 12 millisecond and GCC\_PHAT is used for TDOA estimation.

@@

We also calculated the localization error for each tested point in the region. Figure~\ref{fig:error_distribution} shows a heatmap of the error distribution inside the grid. The error is below $3$ cm for most areas inside the region. There is one error spike in the mid-left region and we contribute this to audio source placecment error becuase the error is fairly low and consistent around that spike region.

@@

Fig~\ref{fig:circle_normal} shows results for experiments with at normal speed, and Fig~\ref{fig:circle_fast} shows results at fast speed. By comparing localization error for each audio source between normal movement speed and fast movement speed, we find the localization error does not depend on how fast the sound sorce is moving. For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.

From fig~\ref{fig:circle_normal}, localization error is $0.9$ cm for white noise, $1.29$ cm for Music A, and $2.9$ cm for Music B. Localization accuracy is the best for white noise source, and worst for Music B. This is consistent with our expectation because low amplitude regions in Music B would cause arrays to lose track of where the source is. This can also be seen from the "blank" regions in fig~\ref{fig:circle_musicbn}.

>>>>

\subsubsection{Movement tracking}
To test how well the arrays track movement, we mounted a rotating disk $40$ centimeter in diameter onto the grid at $(x=0$ cm$, y=-30$ cm$)$. Fig~\ref{fig:setup_circle} shows a picture of the sestup. A sound source is placed on the edge of rotating disk and the arrays localizes the sound source as it rotates in a circle. In this experiment, we tested how accuracy changes with:
\begin{itemize}
\item different window sizes
\item different audio sources
\item different movement tracking filters
\item different movement speeds
\end{itemize}

@@

\item[Music A] A music that has normal audio amplitude throughout experiment period was chosen. ``Honest Eyes'' by Black Tide was the music used.

\item[Music B] A music with intermittent low amplitude sections was chosen. ``Canon'' was the music used.

@@

To test how accuracy varies with window size, the algorithm is fed with recorded microphone data with different segment length. Fig~\ref{fig:accuracy_vs_window} shows how accuracy changes with window length for three GCC algorithms. The error decreases as window size increases and plataeus after window size exceeds around $10$ millisecond. The lowest error achieved is $2.53$ centimeters. It is achieved when window size is set to 12 millisecond and GCC\_PHAT is used for TDOA estimation.

@@

We also calculated the localization error for each tested point in the region. Figure~\ref{fig:error_distribution} shows a heatmap of the error distribution inside the grid. The error is below $3$ cm for most areas inside the region. There is one error spike in the mid-left region and we contribute this to audio source misplacecment becuase the error is fairly low and consistent around that spike region.

@@

Fig~\ref{fig:circle_normal} shows results for experiments at normal speed, and Fig~\ref{fig:circle_fast} shows results at fast speed. By comparing localization error for each audio source between normal movement speed and fast movement speed, we find the localization error does not depend on how fast the sound sorce is moving. For example fig~\ref{fig:circle_musican} and fig~\ref{fig:circle_musicaf} shows that localization error is $1.289$ cm at normal movement speed and $1.291$ cm at fast movement speed.

From fig~\ref{fig:circle_normal}, localization error is $0.9$ cm for white noise, $1.29$ cm for Music A, and $2.9$ cm for Music B. Localization accuracy is the best for white noise source, and worst for Music B. This is consistent with our expectation because low amplitude regions in Music B would cause arrays to lose track of where the source is. This can also be seen from the ``blank'' regions in fig~\ref{fig:circle_musicbn}.

### 2015-05-20 reports/paper/prior.tex
paper update

The requirement of having a template for each location to be detected makes this approach too restrictive for our project, since we want the drawing to be continuous in the 2D region.  Moreover, the need to recalibrate all locations during setup is too curbursome for end users in a portable drawing application. Therefore, our main focus is on TDOA approaches.

@@

Cross-correlation has been used extensively in measuring signal delays\cite{tdoa:mit2, tdoa:3}. Cross-correlation with different prefiterings (known as \emph{generalized cross correlation (GCC)}) have also been investigated to improve delay estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.

GCC provides a framework to estimate delay differences $t_0$ between two signals $x_1(t)$ and $x_2(t)$:
\begin{eqnarray} \label{eq:gcc}
t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
\end{eqnarray}
, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transform of $x_1(t)$ and $x_2(t)$. $W(\omega)$ provides a way to prefilter signals passed to cross correlation estimator. We focused with three ways of prefiltering the signal:
\begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is normal cross correlation.
\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divideded by its magnitude. Only phase information contributes to delay estimation.
\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. part of magnitude information is included in delay estimation.
\end{description}

>>>>

The requirement of having a template for each location to be detected makes this approach too restrictive for our project, since we want the drawing to be continuous in the 2D region.  Moreover, the need to recalibrate all locations during setup is too curbursome for end users in a portable localization system. Therefore, our main focus is on TDOA approaches.

@@

Cross-correlation has also been used in measuring signal delays\cite{tdoa:mit2, tdoa:micloc, tdoa:3}. Cross-correlation with different prefiterings (known as \emph{generalized cross correlation (GCC)}) have also been investigated to improve delay estimation~\cite{tdoa:gcc1,tdoa:gcc2,tdoa:gcc3}.

GCC provides a framework to estimate delay differences $t_0$ between two signals $x_1(t)$ and $x_2(t)$:
\begin{eqnarray} \label{eq:gcc}
t_0 &=& \arg\max_{\tau} R_{x_1x_2}(\tau) \\\label{eq:gcc2}
R_{x_1x_2}(\tau) &=& \int_{-\infty}^\infty W(\omega) X_1(\omega) X_2^{*}(\omega) e^{j\omega\tau} d\omega
\end{eqnarray}
, where $X_1(\omega)$ and $X_2(\omega)$ are Fourier Transform of $x_1(t)$ and $x_2(t)$. $W(\omega)$ provides a way to prefilter signals passed to cross correlation estimator. We focused on three ways of prefiltering the signal:
\begin{description}[\IEEEsetlabelwidth{Very very long label}\IEEEusemathlabelsep]
\item[GCC] $W(\omega) = 1$. No prefiltering is done. This is unfiltered normal cross correlation.
\item[GCC\_PHAT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^{*}(\omega)\right|}$. Each frequency is divideded by its magnitude. Only phase information contributes to delay estimation.
\item[GCC\_PHAT\_SQRT] $W(\omega) = \frac{1}{\left|X_1(\omega)X_2^*(\omega)\right|^{0.5}}$. This is somewhere between GCC and GCC\_PHAT. part of magnitude information is included in delay estimation.
\end{description}

### 2015-05-20 reports/paper/simulation.tex
paper update

To see how sound source distance affects localization accuracy, the same simulation is carried out with sound source moved from $20$ cm to $80$ cm away from the center of the array. Results are presetned in fig~\ref{fig:sim_3_8}. Comparing with fig~\ref{fig:sim_3_2}, accuracy decreases with distance to the array. This is also consistent with our observation in $2$ microphone case where source farther away would result in larger uncertainty region.

@@

In fig~\ref{fig:sim_hm_3_line}, three microphones are placed $10$ cm apart from each other on x-axis. Error heatmap showed high uncertainty on the x axis, and the overall accuracy is not as good as that with three microphones placed in a triangle. The average error is $55.05$ cm.

To further increase the distance between microphones, we placed four microphones at four corners of the region. Fig~\ref{fig:sim_hm_4} showed the result. With this configuration, accuracy is consistently good across the region. The average error is $0.05$ cm.  However, placing microphones far apart at corners of the region requires accurate placement of all four individual microphones. The system is less portable compared to small arrays with microphones near each other.

To avoid the need to accurate place four microphones at far distance(as required by fig~\ref{fig:sim_hm_4}), we explored configuration with two arrays. Two $3$ microphone array are placed $1$ meter apart and the result is presented in fig~\ref{fig:sim_hm_2_array}.  The result indicates that this configuration has good accuracy when source is close to the arrays. Accuracy decreases as sound source moves outside the one meter by one meter region. The average error is $2.60$ cm.

>>>>

To see how sound source distance affects localization accuracy, the same simulation is carried out with the sound source moved from $20$ cm to $80$ cm away from the center of the array. Results are presetned in fig~\ref{fig:sim_3_8}. Comparing with fig~\ref{fig:sim_3_2}, accuracy decreases with distance to the array. This is also consistent with our observation in $2$ microphone case where source farther away would result in larger uncertainty region.

@@

In fig~\ref{fig:sim_hm_3_line}, three microphones are placed $10$ cm apart from each other on x-axis. Error heatmap shows high uncertainty on the x axis, and the overall accuracy is not as good as that with three microphones placed in a triangle. The average error is $55.05$ cm.

To further increase the distance between microphones, we placed four microphones at four corners of the region. Fig~\ref{fig:sim_hm_4} shows the result. With this configuration, accuracy is consistently good across the region. The average error is $0.05$ cm.  However, placing microphones far apart at corners of the region requires accurate placement of all four individual microphones. The system is less portable compared to small arrays with microphones near each other. Placing microphones far apart from each other also causes problems in TDOA estimation, because sampling of microphones in the same array requires synchronized clock.

To avoid the need to accurately place four microphones at far distances(as required by fig~\ref{fig:sim_hm_4}), we explored configuration with two arrays. Two $3$ microphone array are placed $1$ meter apart and the result is presented in fig~\ref{fig:sim_hm_2_array}.  The result indicates that this configuration has good accuracy when source is close to the arrays. Accuracy decreases as sound source moves outside the one meter by one meter region. The average error is $2.60$ cm.

### 2015-05-20 reports/paper/system.tex
paper update

To speed up computation for real time localization, instead of searching for $t_0$ that maximizes equation~\ref{eq:gcc}, a grid search in 2D grid is performed. For each point in the grid, the theoretical TDOA to each microphone pair can be precomputed. Then localization resolves to calculating GCC for each microphone pair and performing a lookup for each point in the grid. To further improve localization accuracy, instead of using point estimate for TDOA, a liklihoood map is built. Each entry in GCC output is used as the liklihood for that delay. With three microphones $m_1,m_2,m_3$, there are three microphones pairs: $m_1m_2,m_1m_3,m_2m_3$. Theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$. Then the liklihood map can be built as:
\begin{eqnarray*}
  L(x,y) &=& R_{m_1,m_2}(D_{m_1,m_2}(x,y)) + R_{m_1,m_3}(D_{m_1,m_3}(x,y)) \\
 & & +R_{m_2,m_3}(D_{m_2,m_3}(x,y)) 
\end{eqnarray*}
,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_3}(\tau)$, and $R_{m_2,m_3}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.

@@

To see the effect of using multiple arrays, fig~\ref{fig:liklihood} shows the individual liklihood map from each array and also the combined liklihood map according to equation~\ref{eq:combine_l}. Individual array gives accurate angle estimate, but has high uncertainty in distance estimate. By combining estimates from two arrays, the angle estimate can be effectively combined to estimate distance.

>>>>

To handle uncertainty in TDOA estimation, instead of using point estimate that maximizes equation~\ref{eq:gcc}, we take the cross-correlation function~\ref{eq:gcc2} as a mapping from delay to liklihood. For each microphone array, we build a liklihood heatmap of each point in the grid being the sound sorce. For each point in the grid, the theoretical TDOA to each microphone pair can be precomputed. Then the heatmap can be generated by going through points in the grid and perform a lookup using equation~\ref{eq:gcc2}. With three microphones $m_1,m_2,m_3$, there are three microphones pairs: $m_1m_2,m_1m_3,m_2m_3$. Theoretical TDOA from each location $(x,y)$ to each microphone pair is precomputed and stored in $D_{m_1,m_2}(x,y)$, $D_{m_1,m_3}(x,y)$, and $D_{m_2,m_3}(x,y)$. Then the liklihood map $L(x,y)$ can be built as:
\begin{eqnarray*}
  L(x,y) &=& R_{m_1,m_2}(D_{m_1,m_2}(x,y)) + R_{m_1,m_3}(D_{m_1,m_3}(x,y)) \\
 & & +R_{m_2,m_3}(D_{m_2,m_3}(x,y)) 
\end{eqnarray*}
,where $R_{m_1,m_2}(\tau)$,$R_{m_1,m_3}(\tau)$, and $R_{m_2,m_3}(\tau)$ denote GCC output from microphone pairs $m_1m_2,m_1m_3,$ and $m_2m_3$.

@@

To see the effect of accuracy improvement using multiple arrays, fig~\ref{fig:liklihood} shows a real life localization where the source is placed at $(0$ cm$,-30$ cm$)$ . It shows the individual liklihood map from each array and also the combined liklihood map according to equation~\ref{eq:combine_l}. Individual array gives accurate angle estimate, but has high uncertainty in source distance. It demonstrated that by combining estimates from two arrays the system is able to perform more accurate localization.
